Can machine learning correct microwave humidity radiances for the influence of clouds?

### 2020-11-26 Article/ici_submm_cloudcorr.tex
version submitted to AMT

The proposed cloud correction scheme makes use of a Quantile Regression Neural
Network (QRNN, \citet{pfreundschuh:aneur:18}) to obtain a probabilistic prediction of the NFCS value. Unlike traditional neural networks techniques, which typically only provide a point
estimate of the target variable, QRNNs are trained to predict an arbitrary set
of quantiles of its Bayesian a posteriori distribution
\citep{pfreundschuh:aneur:18}. The predicted a posteriori distribution 
can then be used to derive an estimate of the NFCS value together with an estimate
of the corresponding uncertainty.

@@

A description of the data used in this study and the QRNN approach is provided
in Sect.\ref{data_methods}. In Sect.~\ref{qrnn_mwhs} we demonstrate the
applicability of correction scheme to existing sensors, and later its
application is extended to include sub-mm channels (Sect.\ref{qrnn_ici}). The
results are discussed in Sect.~\ref{discussions}, and Sect.~\ref{conclusions}
presents the conclusions from this work and the future outlook.

\section{Data and methods}
\label{data_methods}
%
\subsection{Satellite Instruments}
%
\subsubsection{ MicroWave Humidity Sounder-2}
%
%t
\begin{table}[t]
	\caption{Specifications of MWHS-2 channels relevant to this study.}
	\label{tab:specifications_MWHS2}	
	\begin{tabular}{crrr}
		\tophline
		Channel & Frequency 	& Bandwidth & NE$\Delta$T \\
		& [GHz]			& [MHz]		& [K]		\\
		\middlehline
		1	&	89.0   		  & 1500			&	1.0	\\
		6	&	118.75±1.1    & \phantom{0}200 	&	1.6\\
		7	&	118.75±2.5    & \phantom{0}200 	&	1.6\\
		10	&	150.0         & 1500 			&	1.0 \\
		11	&	183.31±1.0      & \phantom{0}500  &	1.0 \\
		12  & 	183.31±1.8    & \phantom{0}700 	&   1.0\\
		13  & 	183.31±3.0      & 1000    		&	1.0	\\
		14  & 	183.31±4.5    & 2000    		&	1.0\\
		15  & 	183.31±7.0      & 2000  			&	1.0  \\
		\bottomhline
	\end{tabular}
	\belowtable{} % Table Footnotes
\end{table}
The MicroWave Humidity Sounder 2 (MWHS-2) is an instrument on two current satellites in the FengYun-3 series, FY-3C and FY-3D. MWHS-2 is a cross track scanning microwave radiometer and measures 15 frequencies in the range 89--191\,GHz. 89\,GHz and 150\,GHz are window channels, five humidity sounding channels are centered around 183\,GHz, and eight temperature sounding channels are centered on the 118.75\,GHz oxygen absorption line. The five humidity sounding channels are similar to ATMS. Observations from MWHS-2 are routinely assimilated in all-sky conditions at the European Centre for Medium-Range Weather Forecasts (ECMWF) with demonstrable positive impact on forecast performance \citep{duncan2020MWHS}. The channels relevant to this study are described in Table~\ref{tab:specifications_MWHS2}.

@@

\subsubsection{Artic Weather Satellite}
%
\begin{table}[t]
	\caption{Specifications of AWS channels relevant to this study.}
	\label{tab:specifications_AWS}	
	\begin{tabular}{lrrr}
		\tophline
		Channel & Frequency 	& Bandwidth & NE$\Delta$T \\
				& [GHz]			& [MHz]		& [K]		\\
		\middlehline
		AWS-32	&	176.31    & 2000	&	0.45	\\
		AWS-33	&	178.81    & 2000 	&	0.45\\
		AWS-34	&	180.31    & 1000 	&	0.64\\
		AWS-35	&	181.51    & 1000 	&	0.64 \\
		AWS-36	&	182.31    & \phantom{0}500  &	0.88 \\
		AWS-41  & 325.15$\pm$6.60    & 2800 	 &0.60\\
		AWS-42  & 325.15$\pm$4.10    & 1800    &0.75	\\
		AWS-43  & 325.15$\pm$2.40    & 1200    &0.92\\
		AWS-44  & 325.15$\pm$1.20    & \phantom{0}800  &1.12  \\
		\bottomhline
	\end{tabular}
	\belowtable{} % Table Footnotes
\end{table}
The AWS is a small satellite mission approved as European Space Agency (ESA) Earth Watch Programme Element. It is a small platform carrying a single across-track scanning microwave radiometer. It is planned as a small but cost-effective platform, which can supplement the information from other polar orbiting satellites. The first prototype mission is expected to be delivered around 2023. The exact channels onboard AWS are not yet set, but five 183\,GHz and four 325\,GHz channels are presently planned. A brief summary of these channels is provided in Table~\ref{tab:specifications_AWS}.

@@

In spite the fact that QRNN only provides a partial cloud correction, the results for  MWHS-2 channel 14 are better than what we achieve with existing cloud filtering techniques like scattering index (SI) and the filtering scheme by \citet{buehler:aclou:07}, hereafter B183. SI uses the differences of brightness temperatures between 89 and 150\,GHz to identify cloud affected data, whilst in B183, they use brightness temperature based threshold to filter out cloudy cases. They recommend a viewing angle dependent brightness temperature threshold at $183.31\pm1.00$\,GHz and brightness temperature difference between $183.31\pm3.00$\,GHz and $183.31\pm 1.00$\,GHz as a measure of cloud impact. The results obtained with SI and B183 are displayed in last two columns of Table~\ref{tab:error_statistics_mwhs_14}. With SI threshold 5\,K, almost 28\,\% of the data is rejected, yet the resulting error distributions are poorer as compared to QRNN. The low bias and skewness values indicate that most of the high negative departures are removed, but cases with low cloud impact pass the filter as clear. Similar is the case with B183. Here only 3\,\% of the data is filtered out but the overall statistics are worse than both QRNN and SI. The results from the two filters are not surprising as both are partial filters and aim at removing only cases with high ice content. The low hydrometeor impact cases remain unaltered.

@@

The calibration of the prediction intervals given out by QRNN is displayed in Fig.~\ref{fig:calibration_mwhs}. We also analyse the calibration of the observational error model used for MWHS-2 in ECMWF NWP system \citep{lawrence2018FY3C}. A gaussian error model is used to represent the distribution of the errors, and plotted under the label ``SI''. For the entire dataset, the predictions from QRNN follow the $y=x$ line, i.e. the predicted uncertainties are perfectly calibrated with the errors observed on the test data. However for the cases with correction greater than 5\,K, the distribution is poorly calibrated, and the curve lies below the $y =x$ line, indicating that the prediction intervals are too narrow. This is in agreement with the wider spread of uncertainties for cases with low accuracy (Fig.~\ref{fig:prediction_uncertainty_mwhs}). However, such cases form less than 6\,\% of the dataset. On the other hand, the calibration of ECMWF error model is above the diagonal for predicted probabilities above 0.2. This suggests that the true probability is higher than what is predicted on these intervals.

@@

\subsection{Prediction accuracy}
\subsubsection{ICI}
%
\begin{figure*}[t]
	\includegraphics[width=\textwidth]{Figures/fig08.pdf} 
	\caption{Same as Fig.~\ref{fig:error_distribution_mwhs14}, but from QRNN-single experiment for ICI channels (a) I1V, (b) I2V and (c) I3V. }
	\label{fig:error_distributions}	
\end{figure*}
\begin{table*}[t]
	\caption{ Same as Table~\ref{tab:error_statistics_mwhs_others}, but from QRNN-single for ICI channels I1V, I2V and I3V. The fraction of rejected cases is given in parentheses.}
	\label{tab:error_statistics_ici}
	%	\tabcolsep=0.11cm
	\begin{tabular}{llrr|rr}
		\tophline
		&&\multicolumn{2}{c|}{Simulations}& \multicolumn{2}{c}{QRNN-single} \\
		\cline{3-6}
		%		\hline
		&&Clear-sky &   All-sky &  \multicolumn{2}{c}{Pred.}  \\
		&&			   &			&	(All) & (5\,K) \\
		\middlehline
		%		\multicolumn{7}{c}{Channel$-$ I1V}\\

@@

\subsection{Prediction uncertainty}
\label{sec:prediction_uncertainty}
%f 
\begin{figure}[t]
	\includegraphics[width = 70mm]{Figures/fig09.pdf}	
	\caption{Same as Fig.~\ref{fig:prediction_uncertainty_mwhs}, but from QRNN-single for ICI channel I2V. The blue line represents a Gaussian distribution with a standard deviation of $0.65$\,K.}
	\label{fig:prediction_uncertainty_I2V}	
\end{figure}
\begin{figure}[t]
	\includegraphics[width = 70mm]{Figures/fig10.pdf}	
	\caption{Same as Fig.~\ref{fig:calibration_mwhs}, but from QRNN-single for ICI channel I2V. }
	\label{fig:calibration_I2V}	
\end{figure}
\begin{figure}[t]
	\includegraphics[width=70mm]{Figures/fig11.pdf}	
	\caption{Same as Fig.~\ref{fig:predicted_errors_mwhs}, but from QRNN-single for ICI channel I2V.}
	\label{fig:predicted_errors}	
\end{figure}

@@

\section{Conclusion and outlook}  %% \conclusions[modified heading if necessary]
\label{conclusions}
%

@@

%% The following commands are for the statements about the availability of data sets and/or software code corresponding to the manuscript.
%% It is strongly recommended to make use of these sections in case data sets and/or software code have been part of your research the article is based on.

@@

\appendix
\section{ARTS setup}    %% Appendix A
\label{appendix:ARTS_setup}
%
All radiative transfer forward simulations are made by the Atmospheric Radiative
Transfer System \citep[ARTS,][]{eriksson:arts2:11,buehler:artst:18}), version
arts-2.3.1. All simulations are based on dBZ-based model system \citep{ekelund2020using}, i.e., CloudSat reflectivities are used as input and are converted to ice water path (IWP) fields using microphysical assumptions. For each atmospheric case, both all-sky and clear-sky calculations are performed. In the former, all hydrometeors contents are set to zero, while in the latter, ice water content (IWC) and rain water content (RWC) derived from CloudSat reflectivities and liquid water content (LWC) from ERA-Interim \citep{dee2011erainterim} are included. In order to avoid a possible bias between clear-sky and all-sky calculations for insignificant hydrometeor contents, both calculations are made by the RT4 solver \citep{evans1995microwavec}.

The absorption model takes into account the effect from nitrogen
\citep{pwr:93}, oxygen \citep{pwr:93} and liquid water content \citep[LWC,][]{ellison2007permittivity}. LWC is taken from ERA-Interim  and is assumed to be totally absorbing. In the mapping of CloudSat reflectivities to RWC and IWC, a total separation between liquid and ice phase is assumed. All scattering hydrometeors at temperatures above 0$^{\degree}$C and below 0$^{\degree}$\,C are assumed to be rain and ice hydrometeors, respectively. For RWC the particle size
distribution (PSD) of \citet{abel2012improved} is applied. The PSD of IWC follows the basic formulation applied in the radio detection and ranging(radar)/light
detection and ranging (lidar) \citep[DARDAR,][\url{
	http://www.icare.univ-lille1.fr/projects/dardar}]{delanoe2008variational}, using latest parameter values (i.e.\ $\alpha$ and $\beta$) as given by \citet{cazenave2019evolution}.
This PSD can be considered as a ``two moment'' scheme, but is here applied in
a one moment manner by setting $N_0^*$ (as a function of temperature)
following Table~5 of \citet{delanoe2014normalized}, and letting the radar
reflectivity set the remaining moment. Single scattering data are taken from
\citet{eriksson:agene:18}. For ice hydrometeors, three habits are applied:
perpendicular 3-bullet rosette, large plate aggregate and large column
aggregate. In the last two cases, the aggregates are complemented with single
crystal data to also cover smaller sizes. These data describe particles
assumed to have a totally random orientation. To apply oriented particles is
much more computationally costly and could not be accommodated inside the
study. The land emissivity was taken from Tool to Estimate Land‐Surface Emissivities at Microwave frequencies \citep[TELSEM,][]{aires2011tool} and the
Ocean/water from Tool to Estimate Sea‐Surface Emissivity from Microwaves to sub‐Millimeter waves \citep[TESSEM,][]{prigent2017sea}.

>>>>

The proposed cloud correction scheme makes use of a Quantile Regression Neural
Network \cite[QRNN,][]{pfreundschuh:aneur:18} to obtain a probabilistic prediction of the NFCS value. Unlike traditional neural networks techniques, which typically only provide a point
estimate of the target variable, QRNNs are trained to predict an arbitrary set
of quantiles of its Bayesian a posteriori distribution
\citep{pfreundschuh:aneur:18}. The predicted a posteriori distribution 
can then be used to derive an estimate of the NFCS value together with an estimate
of the corresponding uncertainty.

@@

A description of the data used in this study and the QRNN approach is provided
in Sect.~\ref{data_methods}. In Sect.~\ref{qrnn_mwhs} we demonstrate the
applicability of correction scheme to existing sensors, and later its
application is extended to include sub-mm channels (Sect.~\ref{qrnn_ici}). The
results are discussed in Sect.~\ref{discussions}, and Sect.~\ref{conclusions}
presents the conclusions from this work, and the future outlook.

\section{Data and methods}
\label{data_methods}
%
\subsection{Satellite Instruments}
%
\subsubsection{ MicroWave Humidity Sounder-2}
%
%t
\begin{table}[t]
	\caption{Specifications of MWHS-2 channels relevant to this study.}
	\label{tab:specifications_MWHS2}	
	\begin{tabular}{crrr}
		\tophline
		Channel & Frequency 	& Bandwidth & NE$\Delta$T \\
		& [GHz]			& [MHz]		& [K]		\\
		\middlehline
		1	&	89.0   		  & 1500			&	1.0	\\
		6	&	118.75±1.1    & \phantom{0}200 	&	1.6\\
		7	&	118.75±2.5    & \phantom{0}200 	&	1.6\\
		10	&	150.0         & 1500 			&	1.0 \\
		11	&	183.31±1.0      & \phantom{0}500  &	1.0 \\
		12  & 	183.31±1.8    & \phantom{0}700 	&   1.0\\
		13  & 	183.31±3.0      & 1000    		&	1.0	\\
		14  & 	183.31±4.5    & 2000    		&	1.0\\
		15  & 	183.31±7.0      & 2000  			&	1.0  \\
		\bottomhline
	\end{tabular}
	\belowtable{} % Table Footnotes
\end{table}
The MicroWave Humidity Sounder 2 (MWHS-2) is an instrument on two current satellites in the FengYun-3 series, FY-3C and FY-3D. MWHS-2 is a cross track scanning microwave radiometer and measures 15 frequencies in the range 89--191\,GHz. 89\,GHz and 150\,GHz are window channels, five humidity sounding channels are centered around 183\,GHz, and eight temperature sounding channels are centered on the 118\,GHz oxygen absorption line. The five humidity sounding channels are similar to ATMS. Observations from MWHS-2 are routinely assimilated in all-sky conditions at the European Centre for Medium-Range Weather Forecasts (ECMWF) with demonstrable positive impact on forecast performance \citep{duncan2020MWHS}. The channels relevant to this study are described in Table~\ref{tab:specifications_MWHS2}.

@@

\subsubsection{Artic Weather Satellite}
%
\begin{table}[t]
	\caption{Specifications of AWS channels relevant to this study.}
	\label{tab:specifications_AWS}	
	\begin{tabular}{lrrr}
		\tophline
		Channel & Frequency 	& Bandwidth & NE$\Delta$T \\
				& [GHz]			& [MHz]		& [K]		\\
		\middlehline
		AWS-32	&	176.31    & 2000	&	0.45	\\
		AWS-33	&	178.81    & 2000 	&	0.45\\
		AWS-34	&	180.31    & 1000 	&	0.64\\
		AWS-35	&	181.51    & 1000 	&	0.64 \\
		AWS-36	&	182.31    & \phantom{0}500  &	0.88 \\
		AWS-41  & 325.15$\pm$6.60    & 2800 	 &0.60\\
		AWS-42  & 325.15$\pm$4.10    & 1800    &0.75	\\
		AWS-43  & 325.15$\pm$2.40    & 1200    &0.92\\
		AWS-44  & 325.15$\pm$1.20    & \phantom{0}800  &1.12  \\
		\bottomhline
	\end{tabular}
	\belowtable{} % Table Footnotes
\end{table}
The AWS is a small satellite mission suggested as an European Space Agency (ESA) Earth Watch Programme Element. It is a small platform carrying a single across-track scanning microwave radiometer. It is planned as a small but cost-effective platform, which can supplement the information from other polar orbiting satellites. The first prototype mission is expected to be delivered around 2023. The exact channels onboard AWS are not yet set, but five 183\,GHz and four 325\,GHz channels are presently planned. A brief summary of these channels is provided in Table~\ref{tab:specifications_AWS}.

@@

In spite the fact that QRNN only provides a partial cloud correction, the results for  MWHS-2 channel 14 are better than what we achieve with existing cloud filtering techniques like scattering index (SI) and the filtering scheme by \citet{buehler:aclou:07}, hereafter B183. SI uses the differences of brightness temperatures between 89 and 150\,GHz to identify cloud affected data; whilst in B183, they recommend a viewing angle dependent brightness temperature threshold at $183.31\pm1.00$\,GHz and brightness temperature difference between $183.31\pm3.00$\,GHz and $183.31\pm 1.00$\,GHz as a measure of cloud impact. The results obtained with SI and B183 are displayed in last two columns of Table~\ref{tab:error_statistics_mwhs_14}. With SI threshold 5\,K, more than 28\,\% of the data is rejected, yet the resulting error distributions are poorer as compared to QRNN. The low bias and skewness values indicate that most of the high negative departures are removed, but cases with low cloud impact pass the filter as clear. Similar is the case with B183. Here only 3\,\% of the data is filtered out but the overall statistics are worse than both QRNN and SI. The results from the two filters are not surprising as both are partial filters and aim at removing only cases with high ice content. The low hydrometeor impact cases remain unaltered.

@@

The calibration of the prediction intervals given out by QRNN is displayed in Fig.~\ref{fig:calibration_mwhs}. We also analyse the calibration of the observational error model used for MWHS-2 in ECMWF NWP system \citep{lawrence2018FY3C}. A gaussian error model is used to represent the distribution of the errors, and plotted under the label ``SI''. For the entire dataset, the predictions from QRNN follow the $y=x$ line, i.e.\ the predicted uncertainties are perfectly calibrated with the errors observed on the test data. However for the cases with correction greater than 5\,K, the distribution is poorly calibrated, and the curve lies below the $y =x$ line, indicating that the prediction intervals are too narrow. This is in agreement with the wider spread of uncertainties for cases with low accuracy (Fig.~\ref{fig:prediction_uncertainty_mwhs}). However, such cases form less than 6\,\% of the dataset. On the other hand, the calibration of ECMWF error model is above the diagonal for predicted probabilities above 0.2. This suggests that the true probability is higher than what is predicted on these intervals.

@@

\subsection{Prediction accuracy}
\subsubsection{ICI}
%
\begin{figure*}[t]
	\includegraphics[width=\textwidth]{Figures/fig08.pdf} 
	\caption{Same as Fig.~\ref{fig:error_distribution_mwhs14}, but from QRNN-single experiment for ICI channels (a) I1V, (b) I2V and (c) I3V. }
	\label{fig:error_distributions}	
\end{figure*}
\begin{table*}[t]
	\caption{ Same as Table~\ref{tab:error_statistics_mwhs_others}, but from QRNN-single for ICI channels I1V, I2V and I3V. The fraction of rejected cases are given in parentheses.}
	\label{tab:error_statistics_ici}
	%	\tabcolsep=0.11cm
	\begin{tabular}{llrr|rr}
		\tophline
		&&\multicolumn{2}{c|}{Simulations}& \multicolumn{2}{c}{QRNN-single} \\
		\cline{3-6}
		%		\hline
		&&Clear-sky &   All-sky &  \multicolumn{2}{c}{Pred.}  \\
		&&			   &			&	(All) & (5\,K) \\
		\middlehline
		%		\multicolumn{7}{c}{Channel$-$ I1V}\\

@@

\subsection{Prediction uncertainty}
\label{sec:prediction_uncertainty}
%f 
\begin{figure}[t]
	\includegraphics[width = 70mm]{Figures/fig09.pdf}	
	\caption{Same as Fig.~\ref{fig:prediction_uncertainty_mwhs}, but from QRNN-single for ICI channel I2V. The blue line represents a Gaussian distribution with a standard deviation of $0.65$\,K.}
	\label{fig:prediction_uncertainty_I2V}	
\end{figure}
\begin{figure}[t]
	\includegraphics[width = 70mm]{Figures/fig10.pdf}	
	\caption{Same as Fig.~\ref{fig:calibration_mwhs}, but from QRNN-single for ICI channel I2V. }
	\label{fig:calibration_I2V}	
\end{figure}
\begin{figure}[t]
	\includegraphics[width= 70mm]{Figures/fig11.pdf}	
	\caption{Same as Fig.~\ref{fig:predicted_errors_mwhs}, but from QRNN-single for ICI channel I2V.}
	\label{fig:predicted_errors}	
\end{figure}

@@

\section{Conclusion and outlook}  
\label{conclusions}
%

@@

@@

\appendix
\section{ARTS setup}    %% Appendix A
\label{appendix:ARTS_setup}
%
All radiative transfer forward simulations are made by the Atmospheric Radiative
Transfer System \citep[ARTS,][]{eriksson:arts2:11,buehler:artst:18}), version
arts-2.3.1. All simulations are based on dBZ-based model system presented by \citet{ekelund2020using}, i.e.\ CloudSat reflectivities are used as input and are converted to ice water content (IWC) and rain water content (RWC) using microphysical assumptions. For each atmospheric case, both all-sky and clear-sky calculations are performed. In the former, all hydrometeors contents are set to zero, while in the latter, IWC and RWC derived from CloudSat reflectivities, and liquid water content (LWC) from ERA-Interim \citep{dee2011erainterim} are included. In order to avoid a possible bias between clear-sky and all-sky calculations for insignificant hydrometeor contents, both calculations are made by the RT4 solver \citep{evans1995microwavec}. The absorption model takes into account the effect from nitrogen
\citep{pwr:93}, oxygen \citep{pwr:93}, liquid water content \citep[LWC,][]{ellison2007permittivity}, and water vapour. For water vapour, absorption model as  in RTTOV \citep[Radiative transfer for TOVS,][]{saunders2018update} is followed, but with few modifications \citep{turner2019amsutran}. LWC is assumed to be totally absorbing. In the mapping of CloudSat reflectivities to RWC and IWC, a total separation between liquid and ice phase is assumed. All scattering hydrometeors at temperatures above 0$^{\degree}$C and below 0$^{\degree}$\,C are assumed to be rain and ice hydrometeors, respectively. For RWC the particle size distribution (PSD) of \citet{abel2012improved} is applied. The PSD of IWC follows the basic formulation applied in DARDAR \citep[\url{
	http://www.icare.univ-lille1.fr/projects/dardar}][]{delanoe2008variational}, using latest parameter values (i.e.\ $\alpha$ and $\beta$) as given by \citet{cazenave2019evolution}.
This PSD can be considered as a ``two moment'' scheme, but is here applied in
a one moment manner by setting $N_0^*$ (as a function of temperature)
following Table~5 of \citet{delanoe2014normalized}, and letting the radar
reflectivity set the remaining moment. Single scattering data are taken from
\citet{eriksson:agene:18}. For ice hydrometeors, three habits are applied:
perpendicular 3-bullet rosette, large plate aggregate and large column
aggregate. In the last two cases, the aggregates are complemented with single
crystal data to also cover smaller sizes. These data describe particles
assumed to have a totally random orientation. To apply oriented particles is
much more computationally costly and could not be accommodated inside the
study. The land emissivity was taken from Tool to Estimate Land‐Surface Emissivities at Microwave frequencies \citep[TELSEM,][]{aires2011tool} and the
Ocean/water from Tool to Estimate Sea‐Surface Emissivity from Microwaves to sub‐Millimeter waves \citep[TESSEM,][]{prigent2017sea}.
